---
title: "A3: SPL Library Checkouts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

# load necessary ALL libraries
library("dplyr")
library("stringr")
library("ggplot2")

# source all R files
source("summary.R")
source("chart_1.R")
source("chart_2.R")
source("chart_3.R")
source("chart_3_2.R")
```

### Introduction

The Seattle Public Library’s Checkouts by Title data allows for viewers to gain a deeper understanding of checkout trends within Seattle’s library system. Providing information about every book or media that has been checked out since 2005, the data can tell users about the popularity and various aspects of books, eBooks, audiobooks, and other forms of physical and electronic media. I chose to focus on one of my favorite childhood series, A Series of Unfortunate Events, written by Lemony Snicket. This series is offered in both eBook and Audiobook form at the Seattle Public Library. Because these books were such a large part of my childhood, I was interested to explore its popularity in this generation of kids. The series has had many renditions during and after its publication which ended in 2006. The series has been retold in both a movie which came out in 2004, as well as a popular series airing on Netflix from 2017-2019. Since the story have managed to stay relevant, let’s see if the books have done the same. 

### Summary Information

```{r echo=FALSE, warning=FALSE, message=FALSE, results="hide", code = readLines("summary.R"), message = FALSE}

```

I looked at the checkout data of the series from January 2022 to January 2023. The total amount of checkouts of every eBook in the series was `r total_checkout_num`, while the total amount of checkouts for the audiobooks was `r total_audio_checkout_num`. The most checked out eBook in the series was `r most_checkouts_ebook` with `r amount_checkouts`, this was the same for the audiobook series with `r amount_audio_checkouts`. Overall the average change in checkout per month for eBooks was `r change_in_checkouts_e`, while the average change per month for audiobooks was `r change_in_checkouts_a`. Overall, this shows us that the audiobooks are generally more popular than the eBooks. Both series’ are pretty consistently checked out each month, not gaining a large amount of readers, but not losing readers either. The first book is by far the most popular for both versions, showing that there is still continual interest in starting the series in today’s readers. 

### The Dataset

The Checkouts by Title dataset was aggregated and published by the Seattle Public Library. This extensive dataset was started in April 2005 and is updated monthly. Because this dataset is so large, I used the data of all checkouts from 2022-2023 to create a smaller dataset for my own project. For this dataset, I chose to focus on Lemony Snicket’s, A Series of Unfortunate Events Book series. The dataset consists of each checked out book’s usage class, checkout type, material type, checkout year, checkout month, checkouts, title, ISBN, creator, subjects, publisher, and publication year. The data was collected from multiple sources, but because this particular series is only offered in eBook and audiobook format, its usage data is provided by OverDrive, hoopla, Freegal, and RBDigital. The Checkouts by Title data was first collected by George Legrady’s art installation in the Central Library. This art piece had been recording checkout data since 2005, and was later used by the Seattle Public Library to create this dataset and share it with Seattle’s open data portal. The dataset was published as part of the Open Data Initiative, urging cities to be transparent with their public data as a means to promote public interaction and problem solving. Because the data is anonymous in both exact checkout date and recipient, there aren’t any ethical implications of publicly sharing this data. My only issue with the dataset was the lack of information regarding how many copies are available of each book. Not knowing their availability makes it hard to compare checkout numbers of different books, since one may be lower simply because there are less copies. This is also misleading when trying to determine the popularity of a book, one with a low number of checkouts may be due to the fact it is constantly checked out and there are only a few versions. I had to base my conclusions off the data given, however, I recognize that this issue may affect the accuracy of my claims. 

### First Trends Over Time Chart

```{r, echo = FALSE, warning = FALSE, code = readLines("chart_1.R"), fig.height = 10, fig.width = 10}
```

This chart shows the total eBook vs audiobook checkouts of the first book in the series, The Bad Beginning, per month from January 2022 to January 2023. I included this chart because how often people are starting the series reflects how popular the books still are. If a lot of people are checking out the first book, it means there is continual interest in the series for those reading for the first time or rereading. From the chart it is clear that the audiobook is much more popular than the eBook. The eBook has its highest number of checkouts in October 2022, with 9 checkouts. The audiobook has the highest number of checkouts in January 2023, with 18 checkouts. Overall, Both versions of the series have continuous checkout trend, with slight variations from month to month. It seems like the audiobook averages around 4-5 checkouts a month, while the audiobook averages around 15-16. This shows that more people are choosing to start the series with the audiobook over the eBook, let’s examine this idea further with the rest of the series.


### Second Trends Over Time Chart

```{r, echo = FALSE, code = readLines("chart_2.R"), fig.height = 10, fig.width = 10}
```

This chart shows the overall checkout count of eBooks vs audiobooks from the series. This chart allows us to understand the overall interaction with each version of the series and determine their overall checkout frequency. From the chart we are able to see that there is more overall interaction with the audiobooks than the eBooks, though both versions get continuous interaction. While audiobook checkouts have continued to remain high, eBook checkouts have been on a constant decline since September 2022. We can see that this was shortly followed by an increase in audiobook checkouts, indicating an increased interest in the audio version over the eBook version. Though eBooks were briefly more popular in May 2022, it is clear that the audiobook version is the preferred way to interact with the series. This confirms the trend we observed in the first chart, people are starting and continuing to choose the audio version of these books. 


### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

```{r, echo = FALSE, code = readLines("chart_3.R"), fig.height = 5, fig.width = 10}

```

```{r, echo = FALSE, code = readLines("chart_3_2.R")}

```



